### 2020-01-29 wuhan.tex
anonymyze papers

\begin{flushleft}{
	\Large
	\textbf\newline{
		Evaluating uncertainties associated with early estimates of the basic reproductive number during the novel coronavirus (2019-nCoV) outbreak
	}
}

@@

First, we gathered information on estimates of $\mathcal R_0$ and their
assumptions about the underlying generation-interval distributions from preprint servers (e.g., bioRxiv, medRxiv, and SSRN) between January 24th, 2020 and January 26th, 2020.
For each study $i$, we model $\mathcal R_0$, $\bar G$, and $\kappa$ with an appropriate probability distribution (either uniform or gamma) that roughly matches the estimated or assumed values and their uncertainty ranges:
\begin{equation}
\begin{aligned}
\mathcal R_{0i} &\sim D_{i, \mathcal R}(\theta_{i, \mathcal R})\\
{\bar G}_i &\sim D_{i, G}(\theta_{i, G})\\
\kappa_i &\sim D_{i, \kappa}(\theta_{i, \kappa})\\
\end{aligned}
\end{equation}
where $D$ is a probability distribution and $\theta$ is a parameter set of the probability distribution $D$.
For example, \cite{imaincov} estimated $\mathcal R_0 = 2.6$ (95\% CI: 1.5--3.5);
we model this parameter as a gamma distribution with mean of 2.6 and a shape parameter of 28, which has a 95\% probability of containing a value between 1.5 and 3.5.
Then, we draw 100,000 random samples from these distributions ($\mathcal R_{0i}^n$, ${\bar G}_i^n$, and $\kappa_i^n$ where $n = 1, \dots, 100,000$) and calculate the exponential growth rate $r$ via the inverse of \eref{gamma}:
\begin{equation}
r_i^n = \frac{{\mathcal R_{0i}^n}^{\kappa_i^n} - 1}{\kappa_i^n \bar{G}_i^n}.
\end{equation}
This allows us to approximte the probability distribution of the exponential growth rate as a discrete probability distribution.

To define appropriate ranges of uncertainties for each parameter ($r$, $\bar G$, and $\kappa$),
we use a Bayesian multilevel modeling approach by assuming that all \emph{observed} (i.e., estimated or assumed) parameter values come from the same distribution:
\begin{equation}
\begin{aligned}
r_i^n &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_r, \mathrm{shape}=\alpha_r),\\
\bar{G}_i^n &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
\kappa_i^n &\sim \mathrm{Exponential}(\mathrm{mean}=\mu_\kappa),\\
(\mu_r, \alpha_r, \mu_G, \alpha_G, \mu_\kappa) &\sim \mathrm{Gamma}(\mathrm{shape}=0.1, \mathrm{rate}=0.1),
\end{aligned}
\end{equation}
where $n$ is randomly selected between 1 and 100,000 at each Metropolis-Hastings step;
this approach is analogous to Bayesian methods for analyzing phylogenetic data, which can rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
We use an exponential distribution to model $\kappa_i$ to account for the fact that the IDEA model, which \cite{majumderncov} uses, assumes a fixed generation interval ($\kappa = 0$).
We run 4 parallel chains that consist of 100,000 burnin steps and 100,000 sampling steps.
Posterior samples are thinned every 200 steps.
Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all parameters \citep{gelman1992inference}.

@@

To evaluate how the uncertainty in each parameter affects the estimate of $\mathcal R_0$,
we replace the estimated or assumed values of $r$, $\bar G$, and $\kappa$ with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate the basic reproductive number $\mathcal R_0$ (\fref{R0}).
We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (15 out 18).
We estimate  narrower confidence intervals for \cite{liuncov} and \cite{majumderncov} when we include uncertainties in $\kappa$ because they assume narrow generation-interval distributions ($\kappa = 0.2$ and $\kappa=0$, respectively);
when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ becomes less sensitive to the values of $r$ and $\bar G$.
We estimate narrower confidence intervals for \citep{riouncov} because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).

We also find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
For example, recalculating $\mathcal R_0$ for \cite{majumderncov} by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
There are two explanations for this result.
First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
$r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ than with $\mu_G$.
Second, assuming a fixed generation time makes their estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.

@@

Finally, we recalculate $\mathcal R_0$ by accounting for all uncertainties (i.e., using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$) and comapre it with the reported $\mathcal R_0$ estimates.
We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.
This result does not imply that our assumptions are too weak;
we believe that this confidence interval accurately reflects the level of uncertainties present in the information that were available when these models were fitted.
In fact, our results show that incorporating uncertainties in all parameters can give narrower confidence intervals than doing so one at a time because nonlinear effects of parameters could cancel out and give appropriately wide confidence intervals.

@@

>>>>

\begin{flushleft}{
	\Large
	\textbf\newline{
		Evaluating uncertainties associated with preliminary estimates of the basic reproductive number during the novel coronavirus (2019-nCoV) outbreak
	}
}

@@

First, we gathered information on estimates of $\mathcal R_0$ and their
assumptions about the underlying generation-interval distributions from
6 articles that were uploaded to
preprint servers (e.g., bioRxiv, medRxiv, and SSRN) between January 24th, 2020 and January 26th, 2020.
For each study $i$, we model $\mathcal R_0$, $\bar G$, and $\kappa$ with an appropriate probability distribution (either uniform or gamma) that roughly matches the estimated or assumed values and their uncertainty ranges:
\begin{equation}
\begin{aligned}
\mathcal R_{0i} &\sim D_{i, \mathcal R}(\theta_{i, \mathcal R})\\
{\bar G}_i &\sim D_{i, G}(\theta_{i, G})\\
\kappa_i &\sim D_{i, \kappa}(\theta_{i, \kappa})\\
\end{aligned}
\end{equation}
where $D$ is a probability distribution and $\theta$ is a parameter set of the probability distribution $D$.
For example, one study estimated $\mathcal R_0 = 2.6$ (95\% CI: 1.5--3.5);
we model this parameter as a gamma distribution with mean of 2.6 and a shape parameter of 28, which has a 95\% probability of containing a value between 1.5 and 3.5.
Then, we draw 100,000 random samples from these distributions ($\mathcal R_{0i}^n$, ${\bar G}_i^n$, and $\kappa_i^n$ where $n = 1, \dots, 100,000$) and calculate the exponential growth rate $r$ via the inverse of \eref{gamma}:
\begin{equation}
r_i^n = \frac{{\mathcal R_{0i}^n}^{\kappa_i^n} - 1}{\kappa_i^n \bar{G}_i^n}.
\end{equation}
This allows us to approximate the probability distribution of the exponential growth rate as a discrete probability distribution.

To define appropriate ranges of uncertainties for each parameter ($r$, $\bar G$, and $\kappa$),
we use a Bayesian multilevel modeling approach by assuming that all \emph{observed} (i.e., estimated or assumed) parameter values come from the same distribution:
\begin{equation}
\begin{aligned}
r_i^n &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_r, \mathrm{shape}=\alpha_r),\\
\bar{G}_i^n &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
\kappa_i^n &\sim \mathrm{Exponential}(\mathrm{mean}=\mu_\kappa),\\
(\mu_r, \alpha_r, \mu_G, \alpha_G, \mu_\kappa) &\sim \mathrm{Gamma}(\mathrm{shape}=0.1, \mathrm{rate}=0.1),
\end{aligned}
\end{equation}
where $n$ is randomly selected between 1 and 100,000 at each Metropolis-Hastings step;
this approach is analogous to Bayesian methods for analyzing phylogenetic data, which can rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
We use an exponential distribution to model $\kappa_i$ since one model assumes a fixed generation interval ($\kappa = 0$).
We run 4 parallel chains that consist of 100,000 burnin steps and 100,000 sampling steps.
Posterior samples are thinned every 200 steps.
Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all parameters \citep{gelman1992inference}.

@@

To evaluate how the uncertainty in each parameter affects the estimate of $\mathcal R_0$,
we replace the estimated or assumed values of $r$, $\bar G$, and $\kappa$ with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate the basic reproductive number $\mathcal R_0$ (\fref{R0}).
We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (15 out 18).
We estimate narrower confidence intervals for Study 2 and Study 3 when we include uncertainties in $\kappa$ because they assume narrow generation-interval distributions ($\kappa = 0.2$ and $\kappa=0$, respectively);
when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ becomes less sensitive to the values of $r$ and $\bar G$.
We estimate narrower confidence intervals when we use the average mean generation time for Study 5 because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).

We also find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
For example, recalculating $\mathcal R_0$ for Study 3 by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
There are two explanations for this result.
First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
$r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ than with $\mu_G$.
Second, assuming a fixed generation time makes their estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.

@@

Finally, we recalculate $\mathcal R_0$ by accounting for all uncertainties (i.e., using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$) and compare it with the reported $\mathcal R_0$ estimates.
We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.
This result does not imply that our assumptions are too weak;
we believe that this confidence interval accurately reflects the level of uncertainties present in the information that were available when these models were fitted.
In fact, our results show that incorporating uncertainties in all parameters can give narrower confidence intervals than doing so one at a time because nonlinear effects of parameters could cancel out and give appropriately wide confidence intervals.

@@

Estimating the basic reproductive number is crucial for predicting the course of an outbreak and planning intervention strategies.
When the basic reproductive numbers are estimated from a wide range of models that rely on different assumptions, 
it can difficult to assess the statistical validity of these estimates.
Here, we present a simple framework for evaluating estimates of the basic reproductive number and associated model assumptions using the gamma approximation of the generation-interval distribution \citep{park2019practical}.
